Reproducibility of COVID-19 pre-prints

### 2021-07-14 outputs/paper/paper-copy.Rmd
Add 2019 arXiv and socArXiv data

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
library(wesanderson)
library(kableExtra)
```

@@

arxiv_open_data_results <- 
  read_csv(here::here("outputs/data/arxiv_open_data_results.csv"))
all_arxiv_results <- 
  read_csv(here::here("outputs/data/arxiv_results.csv")) # All COVID arXiv papers (3,812)

socarxiv_open_data_results <- 
  read_csv(here::here("outputs/data/socarxiv_open_data_results.csv"))
socarxiv_results <- 
  read_csv(here::here("outputs/data/socarxiv_results.csv")) # All COVID medRxiv papers (444). Same as socarxiv_sample.

@@

We find that of the papers sampled, approximately 76 per cent of papers from arXiv, 68 per cent of papers from bioRxiv, 80 per cent of papers from medRxiv, and 83 per cent of papers from SocArXiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \@ref(fig:flowchart). Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains an overall lower proportion of papers with these markers. 
<!-- Finally, we briefly examine the presence of open data and code markers in specific types of pre-prints, namely modelling, simulation, and machine learning research. -->

```{r flowchart, fig.cap = "Summary of process and main results", out.width="90%", fig.align = "center"}
library(DiagrammeR)

DiagrammeR::grViz("digraph {

@@

From that collection of all pre-prints related to COVID019 we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, 1,000 from bioRxiv, and 410 from SocArXiv. Broadly, we are unable to find markers of either open data for open code for 2,987 pre-prints or approximately 76 per cent of our sample (Table \@ref(tab:summarycounts-repositories)). Of the remaining pre-prints, 6 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

Once differentiated by repository, we observe that open data and code markers were absent from 68 per cent of the sampled bioRxiv pre-prints, 80 per cent of the sampled medRxiv pre-prints, 76 per cent of the sampled arXiv pre-prints, and 83 per cent of all COVID-19-related SocArXiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \@ref(tab:summarycounts-repositories)). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 20 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are broadly similar to @@mcguinness2021descriptive, who focus on medRxiv and find that 23 per cent describe open data.

@@

Our metadata also contains an indication of whether a pre-print has been published, in the form of a DOI linking to the corresponding published paper. The proportion of pre-prints that have been published varies by repository (Table \@ref(tab:published-sample-summary)). Notably, of all COVID-19-related pre-prints in our dataset, approximately one third of those posted to bioRxiv, and nearly 30 per cent of those posted to medRxiv, were published in peer-reviewed journals. This is high in comparison to the proportions from arXiv and SocArXiv and may reflect the attention that pre-prints of research from the biomedical and life sciences have received throughout the pandemic.

@@

all_2019 <- rbind(bio_2019, med_2019)

# TODO: position = "dodge" not working
all_2019 %>% 
  mutate(month = floor_date(date, "month")) %>% 
  # filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month, fill = type)) + 
  geom_histogram(stat="count", position = "dodge2") +
  labs(title = "Pre-prints posted per month, 2019",
       subtitle = "Total of 31,752 pre-prints on bioRxiv, 913 on medRxiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1"))
```

```{r 2019-open-results}
just_2019_medRxiv <-
  med_2019_open_data_results  %>%
  mutate(date = as.Date(date))  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(type = "medRxiv")

just_2019_bioRxiv <-
  bio_2019_open_data_results  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(date = as.Date(date)) %>%
  mutate(type = "bioRxiv")

all_2019 <- rbind(just_2019_medRxiv, just_2019_bioRxiv)

all_2019 %>%
  group_by(type) %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  ungroup() %>%
  select(is_open_data, is_open_code, n, prop_total) %>%
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in 2019 medRxiv and bioRxiv samples",
        booktabs = TRUE,
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("bioRxiv" = 4, "medRxiv" = 4))
```

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase which may be expected as the repository gained recognition and popularity in the medical research community. The number of papers posted monthly to bioRxiv also saw an overall increase throughout 2019. It is important to note here that due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 19,638 pre-prints have been posted to medRxiv since June 2019, 12,060 (approximately 61 per cent) of which relate to COVID-19.

@@

published_2019 <- rbind(med_published_2019, bio_published_2019)

@@

published_2019_open_status <- rbind(med_2019_published_open_status, bio_2019_published_open_status)

@@

published_2019_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Published = ifelse(published == 0, "No", "Yes"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Published, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published, 2019",
        booktabs = TRUE,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("bioRxiv" = 2, "medRxiv" = 2))
```

@@

Transparency and reproducibility are hallmarks of quality scientific research. Open data and open code contribute to both by allowing the scientific community to more easily verify the authenticity of purported scientific discovery and its supporting evidence. This is especially important in cases where scientific research may quickly and directly impact clinical practice or public policy such as during the COVID-19 pandemic. Among many other impacts on biomedical research, COVID-19 has dramatically increased the popularity of pre-prints from both a production and consumption standpoint. The number of COVID-19 pre-prints posted to medRxiv increased dramatically in the early stages of the pandemic while non-COVID-19 pre-print numbers were largely as expected. The same trends were apparent in abstracts accessed by medRxiv users, where COVID-19 pre-print abstracts were viewed over 15 times more than non-COVID-19 pre-print abstracts [@citeFraser]. For these reasons, it is important to examine open science standards and reproducibility within pre-print repositories.

@@

An important weakness to note is the potential presence of false negatives in indicators of publication in our dataset. @abdill2019 estimate that the false-negative rate may be as high as 37.5 per cent for data pulled from the bioRxiv API, meaning analysis of published papers may represent only a fraction of those that have actually been published. It is unclear to what extent this is the case for medRxiv or what bias may exist in the subset of pre-prints for which publication was detected, as it is likely that this process relies on title-based text matching [@abdill2019]. It is also likely that some of our more recent sampled pre-prints will be published in future which we could not account for at the time of our data collection.

@@

# Classification

Insert details on keyword search used for machine learning/modelling/simulation

>>>>

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
library(wesanderson)
library(kableExtra)
library(DiagrammeR)
```

@@

arxiv_open_data_results <- 
  read_csv(here::here("outputs/data/arxiv_open_data_results.csv"))
all_arxiv_results <- 
  read_csv(here::here("outputs/data/arxiv_results.csv")) # All COVID arXiv papers (3,812)
arxiv_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/arxiv_2019_open_data_results.csv"))
all_arxiv_2019_results <- 
  read_csv(here::here("outputs/control-data/arxiv_data_2019.csv"))

socarxiv_open_data_results <- 
  read_csv(here::here("outputs/data/socarxiv_open_data_results.csv"))
socarxiv_results <- 
  read_csv(here::here("outputs/data/socarxiv_results.csv")) # All COVID medRxiv papers (410). Same as socarxiv_sample.  
socarxiv_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/socarxiv_2019_open_data_results.csv"))
socarxiv_metadata_2019 <- 
  read_csv(here::here("outputs/control-data/socarxiv_metadata_2019.csv"))

@@

We find that of the papers sampled, approximately 76 per cent of papers from arXiv, 68 per cent of papers from bioRxiv, 80 per cent of papers from medRxiv, and 83 per cent of papers from SocArXiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \@ref(fig:flowchart). Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains approximately the same proportion of papers with these markers. 
<!-- Finally, we briefly examine the presence of open data and code markers in specific types of pre-prints, namely modelling, simulation, and machine learning research. -->

```{r flowchart, fig.cap = "Summary of process and main results", out.width="90%", fig.align = "center"}
DiagrammeR::grViz("digraph {

@@

From that collection of all pre-prints related to COVID-19 we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, 1,000 from bioRxiv, and 410 from SocArXiv. Broadly, we are unable to find markers of either open data for open code for 2,987 pre-prints or approximately 76 per cent of our sample (Table \@ref(tab:summarycounts-repositories)). Of the remaining pre-prints, 6 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

Once differentiated by repository, we observe that open data and code markers were absent from 68 per cent of the sampled bioRxiv pre-prints, 80 per cent of the sampled medRxiv pre-prints, 76 per cent of the sampled arXiv pre-prints, and 83 per cent of all COVID-19-related SocArXiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \@ref(tab:summarycounts-repositories)). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 20 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are broadly similar to @mcguinness2021descriptive, who focus on medRxiv and find that 23 per cent describe open data.

@@

Our metadata also contains an indication of whether a pre-print has been published, in the form of a DOI linking to the corresponding published paper. The proportion of pre-prints that have been published varies by repository (Table \@ref(tab:published-sample-summary)). Notably, of all COVID-19-related pre-prints in our dataset, approximately one third of those posted to bioRxiv, and nearly 30 per cent of those posted to medRxiv, were published in peer-reviewed journals. This is high in comparison to the proportions from arXiv and SocArXiv and may reflect the attention that pre-prints concerning research from the biomedical and life sciences have received throughout the pandemic.

@@

socarxiv_2019 <- socarxiv_metadata_2019 %>%
  select(date_created) %>%
  mutate(type = "socArXiv") %>%
  rename(date = date_created)

arxiv_2019 <- all_arxiv_2019_results %>%
  select(submitted) %>%
  mutate(type = "arXiv") %>%
  rename(date = submitted)

all_2019 <- rbind(bio_2019, med_2019, socarxiv_2019, arxiv_2019)

# TODO: position = "dodge" not working
all_2019 %>% 
  mutate(month = floor_date(date, "month")) %>% 
  # filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month, fill = type)) + 
  geom_histogram(stat="count", position = "dodge2") +
  labs(title = "Pre-prints posted per month, 2019",
       subtitle = "Total of 150,018 pre-prints on arXiv, 31,752 on bioRxiv, 913 on medRxiv, 1,614 on socArXiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1"))
```

```{r 2019-open-results}
just_2019_medRxiv <-
  med_2019_open_data_results  %>%
  mutate(date = as.Date(date), type = "medRxiv")  %>%
  select(date, is_open_data, is_open_code, type)

just_2019_bioRxiv <-
  bio_2019_open_data_results  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(date = as.Date(date), type = "bioRxiv")

just_2019_socarxiv <- 
  socarxiv_2019_open_data_results %>%
  select(date_created, is_open_data, is_open_code) %>%
  mutate(date = as.Date(date_created), type = "socArXiv") %>%
  select(date, is_open_data, is_open_code, type)

just_2019_arxiv <-
  arxiv_2019_open_data_results %>%
  select(submitted, is_open_data, is_open_code) %>%
  mutate(date = as.Date(submitted), type = "arXiv") %>%
  select(date, is_open_data, is_open_code, type)

all_2019 <- rbind(just_2019_medRxiv, just_2019_bioRxiv, just_2019_socarxiv, just_2019_arxiv)

all_2019 %>%
  group_by(type) %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  ungroup() %>%
  select(is_open_data, is_open_code, n, prop_total) %>%
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in 2019 sample",
        booktabs = TRUE,
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("arXiv" = 4, "bioRxiv" = 4, "medRxiv" = 4, "socArXiv" = 4))
```

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase which may be expected as the repository gained recognition and popularity in the medical research community. The number of papers posted monthly to bioRxiv also saw an overall increase throughout 2019. It is important to note here that due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 19,280 pre-prints were been posted to medRxiv between June 2019 and May 1, 2021, 12,060 (approximately 63 per cent) of which relate to COVID-19.

@@

socarxiv_published_2019 <- 
  socarxiv_metadata_2019 %>% 
  select(title, date_created, links.doi) %>% 
  mutate(repo = "socArXiv", published = as.numeric(!is.na(links.doi))) %>%
  rename(date = date_created) %>%
  select(title, date, published, repo)

arxiv_published_2019 <-
  all_arxiv_2019_results %>%
  select(title, submitted, doi) %>%
  mutate(repo = "arXiv", published = as.numeric(!is.na(doi))) %>%
  rename(date = submitted) %>%
  select(title, date, published, repo)

published_2019 <- rbind(med_published_2019, bio_published_2019, socarxiv_published_2019, arxiv_published_2019)

@@

socarxiv_2019_published_open_status <- 
  socarxiv_2019_open_data_results %>%
  mutate(published = !is.na(links.doi) %>% as.numeric()) %>%
  select(title, date_created, published, is_open_code, is_open_data) %>%
  rename(date = date_created) %>%
  mutate(condition = is_open_data + is_open_code, repo = "socArXiv")

arxiv_2019_published_open_status <- 
  arxiv_2019_open_data_results %>%
  mutate(published = !is.na(doi) %>% as.numeric()) %>%
  select(title, submitted, published, is_open_code, is_open_data) %>%
  rename(date = submitted) %>%
  mutate(condition = is_open_data + is_open_code, repo = "arXiv")

published_2019_open_status <- rbind(med_2019_published_open_status, bio_2019_published_open_status, socarxiv_2019_published_open_status, arxiv_2019_published_open_status)

@@

published_2019_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Published = ifelse(published == 0, "No", "Yes"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Published, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published, 2019",
        booktabs = TRUE,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("arXiv" = 2, "bioRxiv" = 2, "medRxiv" = 2, "socArXiv" = 2))
```

@@

Transparency and reproducibility are hallmarks of quality scientific research. Open data and open code contribute to both by allowing the scientific community to more easily verify the authenticity of purported scientific discovery and its supporting evidence. This is especially important in cases where scientific research may quickly and directly impact clinical practice or public policy such as during the COVID-19 pandemic. Among many other impacts on the research landscape, COVID-19 has dramatically increased the popularity of pre-prints from both a production and consumption standpoint. The number of COVID-19 pre-prints posted to medRxiv increased dramatically in the early stages of the pandemic while non-COVID-19 pre-print numbers were largely as expected. The same trends were apparent in abstracts accessed by medRxiv users, where COVID-19 pre-print abstracts were viewed over 15 times more than non-COVID-19 pre-print abstracts [@citeFraser]. For these reasons, it is important to examine open science standards and reproducibility within pre-print repositories.

@@

An important weakness to note is the potential presence of false negatives in indicators of publication in our dataset. @abdill2019 estimate that the false-negative rate may be as high as 37.5 per cent for data pulled from the bioRxiv API, meaning analysis of published papers may represent only a fraction of those that have actually been published. It is unclear to what extent this is the case for other repositories or what bias may exist in the subset of pre-prints for which publication was detected, as it is likely that this process relies on title-based text matching [@abdill2019]. It is also likely that some of our more recent sampled pre-prints will be published in future which we could not account for at the time of our data collection.

@@
